### 2017-06-12 referee_process/second_submission/referee_response.tex
adds template for referee response 2

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[aps, 11pt, singlecolumn]{revtex4-1} % Set the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)
\usepackage{natbib}
\bibliographystyle{apsrev}
\usepackage{setspace}

\begin{document}
%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------
\noindent
Dear Editor,
$\,$\newline

\newenvironment{myquotation}{
\begin{quotation}
\itshape
}{ 
\end{quotation}
}

\begin{singlespace}
Editor letter goes here

$\,$\newline
\noindent
Sincerely,

Evan H. Anders \& Benjamin P. Brown

%\end{singlespace}

$\,$
\newline
$\,$
\newline
\noindent
\Large{\textbf{Response to report B:}}\newline$\,$\newline\indent

\begin{myquotation}

The second submission of this paper, after the first round of review,
is SUBSTANTIALLY different from the first time around, and essentially
needs refereeing from scratch as a new paper. The paper does still
essentially examine the heat transport in compressible convection as
as function of superadiabaticity and Rayleigh number (and therefore
varying thermal diffusivity) for fixed density stratification and
Prandtl number via 2D simulations The new paper adjusted the
formulation of the Nusselt number, changing the results substantially,
and added some 3D simulations to the previously all-2D simulations.
The major result is the scalings of the heat transport (measured by
their Nusselt number) with the Rayleigh number bears a remarkable
similarity to standard Boussinesq Rayleigh-Benard convection, at all
the wide range of superadiabaticities simulated. This in itself is an
interesting result, as it somewhat surprising considering that their
fixed background stratification has quite a strong density contrast
($\sim$20) and some of their departures from adiabaticity are large
($\epsilon \sim$ 1 $\rightarrow$ adiabatic index m=0.5).

I think this paper is getting closer to publishable, although I still
find the explanation of the setup of the model in particular somewhat
opaque though, and the interesting results not really addressed in any
detail, as I'll explain below.

I am glad that the authors took my comments on the setup from before
at least partially to heart! I still think things could be clearer.
Most specifically again, is the question of what are the salient
parameters of compressible convection. There are 4, as mentioned
previously, and the authors here have chosen to keep one fixed -- a
measure of the stratification, for which they use $n_\rho$. For the
others, it is nice to cast things in terms of a Rayleigh number and a
Prandtl number since this gels nicely with standard Boussinesq
Rayleigh-Benard convection (RBC). However, then the third parameter,
the superadiabaticity, which the authors call $\epsilon$, is also part of
the definition of the Rayleigh number, and therefore, as it should be,
the Rayleigh parameter is really a definition of the thermal
diffusivity. To the unaware reader more familiar with RBC, this can be
a bit confusing, since there are no independent measures of the
driving (superadiabaticity) and the thermal diffusivity there. It
would really help the reader here to point out the following in
relation to the simulation sets that are performed:

\begin{itemize}
\item At fixed $\epsilon$, varying Ra means that the thermal diffusivity $\chi_t$
scales like 1/$\sqrt{\text{Ra}}$.
\item At fixed Ra, varying $\epsilon$ means that the thermal diffusivity $\chi_t$
scales like $\epsilon$.
\item Since Pr=fixed, viscous diffusivity scales like thermal diffusivity.
\item (and of course, all vary with depth individually)
\end{itemize}
\end{myquotation}
I need to add this to the experiment section.

\begin{myquotation}
Beyond that, I think everything is right. The expression for the
polytopes in terms of the number of density scale heights makes the
notation a little over-complex. There are some disconnects in where
the non-dimensionalisation is performed (they say that ``take R=1'' at
one point and then non-dimensionalise R out again later, for example).
The description of which of the diffusivities or the
conductivity/dynamic viscosity are constant could certainly be
tightened up.
\end{myquotation}
I need to tighten up the description of polytropes, and only talk about
non-dimensionalizing R once.

\begin{myquotation}
With regard to the results:

In light of what I said above, the results in Fig 2. are not
surprising. Higher Ra here at fixed epsilon means lower thermal
diffusivity, and therefore eddies might be expected to retain their
identity longer. The above explanation would help.
\end{myquotation}
Agreed, we hope that the added explanation makes this clearer to the
reader.

\begin{myquotation}
There is quite a bit of mention of “windy” states without any real
technical description of what they are. Please either give more
information (at least something visual to distinguish from non-windy
states) or remove the distraction.
\end{myquotation}
I think it might be a good idea to remove them in order to keep it 
letter-length, but I'll see if I can keep it short \& explain 
relevant pieces.

\begin{myquotation}
Nusselt number: Usually this is a ratio of heat transport in the
turbulent state to that in the conduction state. So therefore, aren’t
the two $F_A$’s different on the top and the bottom? The top has a
modified kappa but the bottom has the original kappa profile?
\end{myquotation}
Nope!  That's our new change.  Maybe make this clearer.

\begin{myquotation}
I think the paragraphs on the Nu vs Ra and the Re/Pe vs Ra are the
meat of the paper! It would really be nice here to know what causes
the difference between 2D and 3D in the Nu plots. The 2/7 law is often
associated with more “windy” states. That seems LESS likely in 3D, so
what is going on? Note also that the sensitive dependence on the exact
roll or other structure seems to imply that the simulation box is too
small. This dependence should not be the case.
\end{myquotation}
Yeah, AR=4, 8, and 16 all return the windy states, so there's some strange
physics here.  I'll try to explain the discrepancy between 2D and 3D, though.

\begin{myquotation}
The Re vs Ra are a bit mysterious at low Ra. I would expect the Re ~
RA$^{1/2}$. Can you explain these alternative scalings or at least why the
expected scaling emerges at large Ra? Are the 2D ones just
over-constrained?
\end{myquotation}
Yes, can explain: we're combining the 1/2 scaling law that we expect from
the Ra changing diffusivity with the 1/4 scaling law of rising Ma in 2D,
which produces the 3/4 law.

\begin{myquotation}
Regarding Fig 4 — are the authors calculating the integrated evolved
density profile to get this number? If the effect is only in the
boundary layer as they mention, I am surprised that the deviation is
so large. This figure needs much more explanation!
\end{myquotation}
We're calculating the horizontally averaged, time-integrated, evolved density
profile and then taking two measures of it.  One is max/min, one is top of domain
/ bottom of domain.  We should probably explain this more and mention that it persists
even at high $\epsilon$ in 3D.

\noindent
This concludes our response to report B.  We thank referee B for this second report.

$\,$
\newline
$\,$
\newline
\noindent
\Large{\textbf{Response to report C:}}\newline$\,$\newline\indent

\begin{myquotation}
This Letter reports results from a numerical study of thermal
convection in a compressible fluid at unity Prandtl number and a range
of initial stratification. The main result is the scaling of the
Nusselt number with the Rayleigh number, supplemented by some
observations such as characteristic changes in the transition between
subsonic and supersonic regimes.

This problem is highly challenging. Ideally, a numerical exploration
should be guided by some theoretical insights or practical
observations. The present paper appears to be ``purely'' numerical
without such guidance, albeit some comparison with the well-known
problem of incompressible Rayleigh--Bernard convection. Nevertheless,
the results are significant and should be published. While I do not
wholeheartedly recommend this paper to PRL, I would not object to its
acceptance.
\end{myquotation}
Is there any way we can bring in some more ``we're guided by X and thus
we expected to get our results''?

\begin{myquotation}
Below is a minor comment that the authors may find useful.

Most parts of the paper is relatively well written, except in the
introduction. Here, there is room for improvement. The introductory
paragraph is virtually empty. I would cite a couple of papers for
"These prior studies" and be specific about both "important insight"
and known "fundamental properties". The final paragraph of the
introduction has failed to fulfill its mission. I would explicitly
state (at least) the most significant result of the paper in this
paragraph.
\end{myquotation}
We need to beef up the intro, as specified here.

$\,$\newline\noindent
This concludes our response to report C.  We thank referee C for this report.

$\,$
\newline
\noindent

\end{singlespace}

\bibliography{../../biblio.bib}
\end{document}